---
title: "Multivariant AIC Selection 05.19.22"
author: "Zach Compton"
date: "5/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=FALSE}
library(nlme)
library(broom.mixed)
library(huxtable)
library(tidymodels)
library(stringr)
library(rms)
library(phytools)
library(geiger)
library(caper)
library(tidyverse)
library(poolr)
```

```{r, echo=F, warning=F, message=F}
modPgls.SEy = function (model, data, corClass = corBrownian, tree, se = NULL, 
                        method = c("REML", "ML"), interval = c(0, 1000), corClassValue=1, sig2e=NULL, ...) 
{
  Call <- match.call()
  corfunc <- corClass
  spp <- rownames(data)
  data <- cbind(data, spp)
  if (is.null(se)) 
    se <- setNames(rep(0, Ntip(tree)), tree$tip.label)[spp]
  else se <- se[spp]
  
  lk <- function(sig2e, data, tree, model, ve, corfunc, spp) {
    tree$edge.length <- tree$edge.length * sig2e
    ii <- sapply(1:Ntip(tree), function(x, e) which(e == 
                                                      x), e = tree$edge[, 2])
    tree$edge.length[ii] <- tree$edge.length[ii] + ve[tree$tip.label]
    vf <- diag(vcv(tree))[spp]
    w <- varFixed(~vf)
    COR <- corfunc(corClassValue, tree, form = ~spp, ...)
    fit <- gls(model, data = cbind(data, vf), correlation = COR, 
               method = method, weights = w)
    -logLik(fit)
  }
  
  if (is.null(sig2e)) {
    fit <- optimize(lk, interval = interval, data = data, tree = tree, 
                    model = model, ve = se^2, corfunc = corfunc, spp = spp)
    sig2e=fit$minimum
  }
  
  tree$edge.length <- tree$edge.length * sig2e
  ii <- sapply(1:Ntip(tree), function(x, e) which(e == x), 
               e = tree$edge[, 2])
  tree$edge.length[ii] <- tree$edge.length[ii] + se[tree$tip.label]^2
  vf <- diag(vcv(tree))[spp]
  w <- varFixed(~vf)
  obj <- gls(model, data = cbind(data, vf), correlation = corfunc(corClassValue, 
                                                                  tree, form = ~spp, ...), weights = w, method = method)
  obj$call <- Call
  obj$sig2e <- sig2e
  obj
}

#Internal function
pglsSEyPagelToOptimizeLambda=function(lambda,model,data,tree,...) {
  -logLik(modPgls.SEy(model=model,data=data,tree=tree,corClassValue=lambda,corClass=corPagel,fixed=T,...)) #Returns -logLikelihood of the pgls.SEy model with lambda fixed to the value of the lambda argument. sig2e will be optimized within modPgls.SEy unless given as an argument here
}

#Function intended for users
pglsSEyPagel=function(model, data, tree, lambdaInterval=c(0,1),...){
  optimizedModel=optimize(pglsSEyPagelToOptimizeLambda,lambdaInterval,model=model,data=data,tree=tree,...) #Optimizes lambda in the lambdaInterval using the pglsSEyPagelToOptimizeLambda function
  return(modPgls.SEy(model=model,data=data,tree=tree,corClass=corPagel,fixed=T,corClassValue=optimizedModel$minimum,...)) #Returns the final model fit
}


```

### Data & Trees
```{r}
Data <- read.csv("min20-2022.05.16.csv")
tree <- read.tree("min20Fixed516.nwk")
```

Cut in the variables that we need for the complete multivariant regression that includes all sigificant variables. Significant variables include Gestation length, adult weight, litter size, and longevity. Pre-remove all NAs. 
```{r}
cutData <- Data[,c(5,9,10,11,13,30,34,38,40,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)
```

Prune the phylogeny to match species.
```{r}
cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]
cat(nrow(cutData),"Species")

```
Set row names as Species name and extract Standard Error as a vector
```{r}
rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
```
All significant variable regression
```{r}
mltv1.neo<-pglsSEyPagel(NeoplasiaPrevalence~log10(adult_weight.g.)+ log10(max_longevity.months.)+
                          log10(litters_year) 
                        ,data=cutData,tree=pruned.tree,se=SE,method = "ML")

```

```{r, echo = F}
mod1 <- tidy(mltv1.neo)
tab1 <-huxtable(mod1)
print(tab1)
```
Grab R^2 for each life history var
```{r}
r.v.mltv1.neo <- summary(mltv1.neo)$corBeta
r.v.wgt <- format(r.v.mltv1.neo[2,1])
r.v.wgt <-signif(as.numeric(r.v.wgt )^2, digits= 2)
cat("Weight:",r.v.wgt)
r.v.long <- format(r.v.mltv1.neo[3,1])
r.v.long<-signif(as.numeric(r.v.long)^2, digits= 2)
cat("Longevity:",r.v.long)
r.v.litter <- format(r.v.mltv1.neo[4,1])
r.v.litter<-signif(as.numeric(r.v.litter)^2, digits= 2)
cat("Litters per Year:",r.v.litter)

```
Combine p values
```{r}
p.v.mltv<-summary(mltv1.neo)$tTable
p.v.mltvwgt<-signif(p.v.mltv[2,4], digits = 2)
p.v.mltvlong<-signif(p.v.mltv[3,4], digits = 2)
p.v.mltvlit<-signif(p.v.mltv[4,4], digits = 2)

pvalues<-c(p.v.mltvwgt,p.v.mltvlong,p.v.mltvlit)

combop<-fisher(pvalues)


cat("Overall P-Value: ", as.numeric(combop[1]))

```
Lambda
```{r}
ld.v.mltv<- summary(mltv1.neo)$modelStruct$corStruct
ld.v.mltv<- signif(ld.v.mltv[1], digits = 2)
cat("Lambda:",ld.v.mltv)

```





Cut for weight and litters
```{r}
cutData <- Data[,c(5,9,10,11,13,30,34,38,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)

cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]

rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
cat(nrow(cutData),"Species")

```

Weight + Litters per year
```{r}
wgtlit<-pglsSEyPagel(NeoplasiaPrevalence~log10(adult_weight.g.)+
                          log10(litters_year),data=cutData,tree=pruned.tree,se=SE,method = "ML")
```

```{r,echo=F}
mod2 <- tidy(wgtlit)
tab2 <-huxtable(mod2)
print(tab2)
```
Grab R^2 for each life history var
```{r}
r.v.wgtlit <- summary(wgtlit)$corBeta 
r.v.wgt <- format(r.v.wgtlit[2,1])
r.v.wgt <-signif(as.numeric(r.v.wgt )^2, digits= 2)
cat("Weight:",r.v.wgt)
r.v.litter <- format(r.v.wgtlit[3,1])
r.v.litter<-signif(as.numeric(r.v.litter)^2, digits= 2)
cat("Litter per Year:",r.v.litter)

```
Combine p values
```{r}
p.v.wgtlit<-summary(wgtlit)$tTable
p.v.wgtlitwgt<-signif(p.v.wgtlit[2,4], digits = 2)
p.v.wgtlitlit<-signif(p.v.wgtlit[3,4], digits = 2)


pvalues<-c(p.v.wgtlitwgt,p.v.wgtlitlit)

combopwgtlit<-fisher(pvalues)


cat("Overall P-Value: ", as.numeric(combopwgtlit[1]))

```

Lambda
```{r}
ld.v.wgtlit<- summary(wgtlit)$modelStruct$corStruct
ld.v.wgtlit<- signif(ld.v.wgtlit[1], digits = 2)
cat("Lambda:",ld.v.wgtlit)

```

Recut Data for adult weight +long
```{r}
cutData <- Data[,c(5,9,10,11,13,38,40,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)

cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]

rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
cat(nrow(cutData),"Species")
```

wgt+long
```{r}
wgtlong<-pglsSEyPagel(NeoplasiaPrevalence~log10(adult_weight.g.)+log10(max_longevity.months.)
,data=cutData,tree=pruned.tree,se=SE,method = "ML")
coef(wgtlong)
```

```{r, echo = F}
mod4 <- tidy(wgtlong)
tab4 <-huxtable(mod4)
print(tab4)
```
Grab R^2 for each life history var
```{r}
r.v.wgtlong <- summary(wgtlong)$corBeta 
r.v.wgt <- format(r.v.wgtlong[2,1])
r.v.wgt <-signif(as.numeric(r.v.wgt )^2, digits= 2)
cat("Weight:",r.v.wgt)
r.v.long <- format(r.v.wgtlong[3,1])
r.v.long<-signif(as.numeric(r.v.long)^2, digits= 2)
cat("Maximum Longevity:",r.v.long)

```
Combine p values
```{r}
p.v.wgtlong<-summary(wgtlong)$tTable
p.v.wgtlongwgt<-signif(p.v.wgtlong[2,4], digits = 2)
p.v.wgtlonglong<-signif(p.v.wgtlong[3,4], digits = 2)


pvalues<-c(p.v.wgtlongwgt,p.v.wgtlonglong)

combopwgtlong<-fisher(pvalues)


cat("Overall P-Value: ", as.numeric(combopwgtlong[1]))

```

Lambda
```{r}
ld.v.wgtlong<- summary(wgtlong)$modelStruct$corStruct
ld.v.wgtlong<- signif(ld.v.wgtlong[1], digits = 2)
cat("Lambda:",ld.v.wgtlong)

```

Recut Data for adult weight +long mal
```{r}
cutData <- Data[,c(5,9,10,11,17,38,40,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)

cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]

rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
cat(nrow(cutData),"Species")
```

wgt+long
```{r}
wgtlong<-pglsSEyPagel(MalignancyPrevalence~log10(adult_weight.g.)+log10(max_longevity.months.)
,data=cutData,tree=pruned.tree,se=SE,method = "ML")
coef(wgtlong)
```

```{r, echo = F}
mod4 <- tidy(wgtlong)
tab4 <-huxtable(mod4)
print(tab4)
```
Grab R^2 for each life history var
```{r}
r.v.wgtlong <- summary(wgtlong)$corBeta 
r.v.wgt <- format(r.v.wgtlong[2,1])
r.v.wgt <-signif(as.numeric(r.v.wgt )^2, digits= 2)
cat("Weight:",r.v.wgt)
r.v.long <- format(r.v.wgtlong[3,1])
r.v.long<-signif(as.numeric(r.v.long)^2, digits= 2)
cat("Maximum Longevity:",r.v.long)

```
Combine p values
```{r}
p.v.wgtlong<-summary(wgtlong)$tTable
p.v.wgtlongwgt<-signif(p.v.wgtlong[2,4], digits = 2)
p.v.wgtlonglong<-signif(p.v.wgtlong[3,4], digits = 2)


pvalues<-c(p.v.wgtlongwgt,p.v.wgtlonglong)

combopwgtlong<-fisher(pvalues)


cat("Overall P-Value: ", as.numeric(combopwgtlong[1]))

```

Lambda
```{r}
ld.v.wgtlong<- summary(wgtlong)$modelStruct$corStruct
ld.v.wgtlong<- signif(ld.v.wgtlong[1], digits = 2)
cat("Lambda:",ld.v.wgtlong)

```


Recut Data for adult weight
```{r}
cutData <- Data[,c(5,9,10,11,13,38,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)

cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]

rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
cat(nrow(cutData),"Species")
```


Weight
```{r}
wgt<-pglsSEyPagel(NeoplasiaPrevalence~log10(adult_weight.g.)
,data=cutData,tree=pruned.tree,se=SE,method = "ML")
```

```{r, echo = F}
mod4 <- tidy(wgt)
tab4 <-huxtable(mod4)
print(tab4)
```
Grab R^2 

```{r}
r.v.wgt <- summary(wgt)$corBeta 
r.v.wgt <- format(r.v.wgt[2,1])
r.v.wgt <-signif(as.numeric(r.v.wgt )^2, digits= 2)
cat("Weight:",r.v.wgt)
```

Lambda
```{r}
ld.v.wgtlong<- summary(wgtlong)$modelStruct$corStruct
ld.v.wgtlong<- signif(ld.v.wgtlong[1], digits = 2)
cat("Lambda:",ld.v.wgtlong)

```

Recut Data for adult weight + gest
```{r}
cutData <- Data[,c(5,9,10,11,13,38,30,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)

cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]

rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
cat(nrow(cutData),"Species")
```


Weight + gestation
```{r}
wgtgest<-pglsSEyPagel(NeoplasiaPrevalence~log10(adult_weight.g.) +log10(Gestation.months.)
,data=cutData,tree=pruned.tree,se=SE,method = "ML")
```

```{r, echo = F}
mod4 <- tidy(wgtgest)
tab4 <-huxtable(mod4)
print(tab4)
```
Grab R^2 for each life history var
```{r}
r.v.wgtgest <- summary(wgtgest)$corBeta 
r.v.wgt <- format(r.v.wgtgest[2,1])
r.v.wgt <-signif(as.numeric(r.v.wgt )^2, digits= 2)
cat("Weight:",r.v.wgt)
r.v.gest<- format(r.v.wgtgest[3,1])
r.v.gest<-signif(as.numeric(r.v.gest)^2, digits= 2)
cat("Gestation:",r.v.gest)

```
Combine p values
```{r}
p.v.wgtgest<-summary(wgtgest)$tTable
p.v.wgtlongwgt<-signif(p.v.wgtgest[2,4], digits = 2)
p.v.wgtlonggest<-signif(p.v.wgtgest[3,4], digits = 2)


pvalues<-c(p.v.wgtlongwgt,p.v.wgtlonglong)

combopwgtlong<-fisher(pvalues)


cat("Overall P-Value: ", as.numeric(combopwgtlong[1]))

```

Lambda
```{r}
ld.v.wgtgest<- summary(wgtgest)$modelStruct$corStruct
ld.v.wgtgest<- signif(ld.v.wgtgest[1], digits = 2)
cat("Lambda:",ld.v.wgtgest)

```


Recut Data for long + gest
```{r}
cutData <- Data[,c(5,9,10,11,13,40,30,42),drop=FALSE]
cutData[cutData < 0] <-NA
cutData <- na.omit(cutData)
head(cutData)

cutData$Species <- gsub(" ", "_", cutData$Species)
includedSpecies <- cutData$Species
pruned.tree<-drop.tip(
  tree, setdiff(
    tree$tip.label, includedSpecies))
pruned.tree <- keep.tip(pruned.tree,pruned.tree$tip.label)
cutData$Keep <- cutData$Species %in% pruned.tree$tip.label
cutData <- cutData[!(cutData$Keep==FALSE),]

rownames(cutData)<-cutData$Species
SE<-setNames(cutData$SE_simple,cutData$Species)[rownames(cutData)]
cat(nrow(cutData),"Species")
```


gest+long
```{r}
gestlong<-pglsSEyPagel(NeoplasiaPrevalence~log10(max_longevity.months.) +log10(Gestation.months.)
,data=cutData,tree=pruned.tree,se=SE,method = "ML")
```

```{r, echo = F}
mod4 <- tidy(gestlong)
tab4 <-huxtable(mod4)
print(tab4)
```
Grab R^2 for each life history var
```{r}
r.v.gestlong <- summary(gestlong)$corBeta 
r.v.long <- format(r.v.gestlong[2,1])
r.v.long <-signif(as.numeric(r.v.long )^2, digits= 2)
cat("Longevity:",r.v.long)
r.v.gest<- format(r.v.gestlong[3,1])
r.v.gest<-signif(as.numeric(r.v.gest)^2, digits= 2)
cat("Gestation:",r.v.gest)

```
Combine p values
```{r}
p.v.gestlong<-summary(gestlong)$tTable
p.v.gestlonglong<-signif(p.v.gestlong[2,4], digits = 2)
p.v.gestlonggest<-signif(p.v.wgtgest[3,4], digits = 2)


pvalues<-c(p.v.gestlonglong,p.v.gestlonggest)

combopgestlong<-fisher(pvalues)


cat("Overall P-Value: ", as.numeric(combopwgtlong[1]))

```

Lambda
```{r}
ld.v.gestlong<- summary(gestlong)$modelStruct$corStruct
ld.v.gestlong<- signif(ld.v.gestlong[1], digits = 2)
cat("Lambda:",ld.v.gestlong)

```
## AIC
```{r}
aicScores <- data.frame("All Variables" = AIC(mltv1.neo), "Adult weight + Litters per Year"=AIC(wgtlit), 
            "Adult weight + Longevity" =AIC(wgtlong), 
              "Adult weight" = AIC(wgt))
aicScores
```

